### 2019-09-02 Final Submission/SinG19_Auto_CoverLetterRevision.tex

\documentclass[11pt]{letter}
\usepackage{vtmathlet}
%\input{vtmathlet.tex}
\usepackage{amsfonts}
\usepackage{color}
\usepackage{soul}
 \textheight 20.5cm
 \textwidth 16.0cm
% \topmargin -3.0cm
%\evensidemargin 1.0cm
 \oddsidemargin -0.1cm

 \begin{document}

\begin{letter} {Professor Richard Middleton\\
Editor (Systems and Control Theory) \\
Automatica
 }

 \signature{ 
 %Christopher Beattie, 
Klajdi Sinani \\
% and Volker Mehrmann
}

 \opening{Dear  Prof. Middleton:}

 Thank you very much for handling our manuscript ``$\mathcal{H}_2(t_f)$ Optimality Conditions for a Finite-time Horizon"  with submission number 18-0764, submitted to Automatica for a possible publication.

We have revised our manuscript following the referees' comments. In addition to the revised manuscript, we are also including a detailed response explaining our revisions and our answers to the referees' questions and comments.

We have uploaded the revised version, in which the revised parts are highlighted in blue so that the referees can easily spot the changes.

Moreover, following your suggestions, we put the paper in two-column format and now the length is around 10.5 pages.

Looking forward to hearing back from you.

\vspace{1.0cm}

\closing{Sincerely yours,}

\end{letter}

\end{document}